### 2024-02-29 docs/papers/sto/so.tex
FIXUP draft update T and r plots

\documentclass[modern, trackchanges, dvipsnames]{aastex631}
  \urlstyle{sf}

\usepackage{CJK}
\usepackage{microtype}
\usepackage{savesym}
  \savesymbol{lambdabar}
\usepackage{newpxtext, eulerpx}
  \restoresymbol{newpx}{lambdabar}
\usepackage[T1]{fontenc}
\usepackage{fontawesome}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{nicefrac}
\usepackage[caption=false]{subfig}
\usepackage[figure,figure*]{hypcap}
%\usepackage{tikz}
%  \usetikzlibrary{positioning, fit, calc, arrows.meta}
\usepackage{booktabs}

\newcommand{\pmwd}{{\usefont{T1}{nova}{m}{sl}pmwd}}
\newcommand{\GADGET}{{{\fontsize{10pt}{12pt}\selectfont GADGET}-4}}
\newcommand{\GSDATA}{{GS512}}

%\renewcommand{\sectionautorefname}{Sec.}
%\renewcommand{\subsectionautorefname}{Sec.}
%\renewcommand{\appendixautorefname}{App.}
%\renewcommand{\figureautorefname}{Fig.}
%\newcommand{\subfigureautorefname}{\figureautorefname}

%FIXME copied from ../adjoint/adjoint.tex
\newcommand{\deltaD}{\delta^\textsc{d}}
\newcommand{\deltaK}{\delta^\textsc{k}}
\renewcommand{\d}{d}
\newcommand{\p}{\partial}
\newcommand{\cJ}{\mathcal{J}}
\newcommand{\cR}{\mathcal{R}}
\newcommand{\cL}{\mathcal{L}}
\newcommand{\cH}{\mathcal{H}}
\bmdefine{\vzero}{0}
\bmdefine{\vI}{I}
\bmdefine{\vnabla}{\nabla}
\bmdefine{\vtheta}{\theta}  % parameters
\bmdefine{\vvartheta}{\vartheta}  % relevant scales and factors to form
                                  % dimensionless features as NN inputs
\bmdefine{\vomega}{\omega}  % white noise modes
\bmdefine{\vk}{k}  % wavevectors
\bmdefine{\vx}{x}  % comoving and canonical coordinates
\bmdefine{\vq}{q}  % Lagrangian coordinates
\bmdefine{\vs}{s}  % displacements
\bmdefine{\vp}{p}  % canonical momenta
\bmdefine{\vv}{v}  % FIXME some velocities
\bmdefine{\va}{a}  % accelerations
\bmdefine{\vz}{z}  % states
\bmdefine{\vo}{o}  % observable
\bmdefine{\vf}{f}
\bmdefine{\vF}{F}
\bmdefine{\vDelta}{\Delta}
\bmdefine{\vLambda}{\Lambda}
\bmdefine{\vlambda}{\lambda}
\bmdefine{\vvarphi}{\varphi}
\bmdefine{\vxi}{\xi}  % x adjoint
\bmdefine{\vpi}{\pi}  % p adjoint
\bmdefine{\valpha}{\alpha}  % force vjp x gradient
\bmdefine{\vzeta}{\zeta}  % force vjp theta gradient
\newcommand{\lna}{\ln\!a}
\newcommand{\lnk}{\ln\!k}
\newcommand{\half}{\nicefrac12}
\newcommand{\As}{A_\mathrm{s}}
\newcommand{\ns}{n_\mathrm{s}}
\newcommand{\Omegam}{\Omega_\mathrm{m}}
\newcommand{\Omegab}{\Omega_\mathrm{b}}
\newcommand{\OmegaK}{\Omega_K}
\newcommand{\Mpc}{\mathrm{Mpc}}
\newcommand{\ic}{\mathrm{i}}
\newcommand{\f}{\mathrm{f}}
\newcommand{\linear}{\mathrm{lin}}
\newcommand{\tophat}{\mathrm{TH}}
\newcommand{\gauss}{\mathrm{G}}

\newcommand{\GPU}{NVIDIA H100 PCIe}  % 80GB memory

\newcommand{\YL}[1]{\textcolor{Bittersweet}{#1}}
\newcommand{\YZ}[1]{\textcolor{MidnightBlue}{#1}}

\begin{document}

\title{\large Neural and Symbolic Optimization of Cosmological
Particle-Mesh Simulation
\vspace{0.3em}}

\author[0000-0002-9300-2632]{\normalsize Yucheng Zhang}
\affiliation{Peng Cheng Laboratory, Shenzhen, Guangdong 518066, China}
\affiliation{Center for Cosmology and Particle Physics, New York
University, New York, New York 10003, USA}

\author[0000-0002-0701-1410]{\normalsize Yin Li}
\affiliation{Center for Computational Mathematics, Flatiron Institute,
New York, New York 10010, USA}
\affiliation{Peng Cheng Laboratory, Shenzhen, Guangdong 518066, China}
\affiliation{Center for Computational Astrophysics, Flatiron Institute,
New York, New York 10010, USA}

\author[0000-0001-5044-7204]{\normalsize Drew Jamieson}
\affiliation{Max Planck Institute for Astrophysics, 85748 Garching bei
M\"unchen, Germany}

\author[0000-0003-0745-9431]{\normalsize Libin Lu}
\affiliation{Center for Computational Mathematics, Flatiron Institute,
New York, New York 10010, USA}

\shorttitle{Optimized Cosmological Simulation}
\shortauthors{Zhang \& Li et al.}

\correspondingauthor{Yucheng Zhang \& Yin Li}
\email{yucheng.zhang@nyu.edu \& eelregit@gmail.com}

\begin{abstract}
% up to 150 words

We optimize the spatial resolution of cosmological particle-mesh (PM)
simulations by sharpening the PM force with first neural networks (NNs)
and then symbolic regression (SR), taking symmetry and dimensional
analysis in consideration.
We embed deep NNs in the differentiable cosmological $N$-body
simulation, and training them to learn for the force kernel.
We fit the trained NNs with SR to extract an analytic kernel for better
interpretability and physical insights, which is also applicable to all
other cosmological PM simulation code.

\end{abstract}

% Main text: up to 3500 words: introduction + results + discussion
% so we should keep introduction and discussion brief
\vspace{1em}
\section{Introduction}
% without heading

% AI for science and cosmology
Rapid advances in artificial intelligence (AI) and more specifically
deep learning (DL) have provided powerful tools that are revolutionizing
scientific research across disciplines.
AI for science consists of not only the state-of-art deep neural
networks, but also more fundamental computational tools and hardwares
like automatic differentiation (AD) and GPUs.
One particular application is the numerical simulation of the Universe.

% a brief review of (differentiable) cosmological simulations
BORG, ELUCID, and BORG-PM \citep{BORG, ELUCID, BORG-PM}
\texttt{FastPM} and \texttt{FlowPM} \citep{FastPM, vmad, SeljakEtAl2017, FlowPM}

% introduce PM and its limitation on small scales
The particle-mesh (PM) method solves the Poisson equation by scattering
the particles onto a mesh, on which FFT is performed and the Laplacian
operator reduces to simple multiplication of wavenumbers in Fourier
space.
Although being computationally efficient, the PM method is mainly
limited by its resolution on small scales, i.e. scales around the mesh
cell size.
This is due to the loss of information during the scattering process of
continuously distributed particles onto the discrete mesh.
We use the cloud-in-cell (CIC) scheme \citep{HockneyEastwood1988} to
compute the mass fractions of a particle scattered to nearby grid
points.
Other schemes include the nearest grid point (NGP) and triangle shaped
cloud (TSC) etc.

% motivation of this work
Optimizing the accuracy of fast and differentiable simulations is
crucial for their applications in, e.g. generation of mocks for
cosmological surveys, simulation-based inference or field-level
inference.

% introduction of this work
The differentiable cosmological simulation with adjoint method
\citep{Li2022a} and the corresponding PM $N$-body library \pmwd\
\citep{Li2022b} lay the foundation of the gradient-based optimization
discussed in this work.
Within this fully differentiable framework, we sharpen the gravitational
force by embedding neural networks (NNs) in the Poisson equation and
train them for the spatial optimization (SO) of the simulation
resolution on small scales.
For better interpretability, we further fit the trained NNs using
symbolic regression (SR).

\vspace{1em}
\section{Results}
% Results should be divided by topical subheadings
% Put main ideas in Results first, then more technical details in Methods.

\subsection{Optimization of Dynamics}

\begin{table*}
  \centering
  \caption{Ranges of \GADGET\ and \pmwd\ configurations and cosmological
  parameters.
  Note that the grid ratio need next fast len to determine the mesh shape
  for fast FFT.
  Given the box size, the mesh shape determines the cell size.
  The softening parameter gives the ratio of the comoving softening length
  to the mean particle spacing.
  The curvature $\OmegaK$ is related to the separate universe simulation.
  We sample parameters applicable to \GADGET\ during data generation, and
  those applicable to \pmwd\ during training.
  \YL{(I forgot why I wanted to sample mesh shape and mesh cell size...
  Now it seems more natural to me to have the ptcl spacing sampled
  uniformly instead...)}
  }
  \label{tab:param}
  \begin{tabular}{lcccr}
  \toprule
  applicability & parameter & distribution & range & sampling \\
  \midrule
  & $128^3$ white noise & $\mathcal{N}(\vzero, \vI)$ & 512 realizations & MC \\
  \cmidrule(lr){2-5}
  & box size in Mpc$/h$ & \YL{log-uniform} & [25.6, 2560) \\
  \cmidrule(lr){2-4}
  & snapshot offset & uniform & (0, 1] \\
  \cmidrule(lr){2-4}
  \GADGET\ & $\As \times 10^9$ & log-uniform & [1, 4) \\
  \cmidrule(lr){2-4}
  \& \pmwd\ & $\ns$ & log-uniform & [0.75, 1.25) & RQMC \\
  \cmidrule(lr){2-4}
  & $\Omegam$ & log-uniform & [1/5, 1/2) & see \autoref{fig:sobol} \\
  \cmidrule(lr){2-4}
  & $\Omegab / \Omegam$ & log-uniform & [1/8, 1/4) \\
  \cmidrule(lr){2-4}
  & $\OmegaK / (1 - \OmegaK)$ & uniform & [-1/3, 1/3) \\
  \cmidrule(lr){2-4}
  & $h$ & log-uniform & [0.5, 1) \\
  \cmidrule(lr){1-4}
  \GADGET\ & softening ratio & log-uniform & [1/50, 1/20) \\
  \cmidrule(lr){1-5}
  \pmwd\ & mesh shape & log-uniform & [128, 512] & MC \\
  \cmidrule(lr){2-4}
  \cmidrule(lr){1-4}
  training & loss grid offset & uniform & $[0, 1)^3$ \\
  \bottomrule
  \end{tabular}
  \end{table*}

\begin{table}
  \centering
  \caption{Possible features that can affect force sharpening.
  All scales are then multiplied by the wavenumber $k$ to form dimensionless
  factors.}
  \label{tab:feat}
  \begin{tabular}{rl}
  \toprule
  type & feature \\
  \midrule
  & $R_\mathrm{d}$: RMS linear displacement \\
  \cmidrule(lr){2-2}
  & $R_P$: $\Delta_\linear^2(R_P^{-1}) \triangleq 1$ \\
  \cmidrule(lr){2-2}
  nonlinear scales & $R_\tophat$: $\sigma_\tophat^2(R_\tophat) \triangleq 1$ \\
  \cmidrule(lr){2-2}
  & $R_\gauss$: $\sigma_\gauss^2(R_\gauss) \triangleq 1$ \\
  \cmidrule(lr){2-2}
  & $\d R / \d\lna$: time derivatives of the above \\
  \midrule
  & particle spacing \\
  \cmidrule(lr){2-2}
  other scales & $l$: PM force mesh cell size \\
  \cmidrule(lr){2-2}
  & softening length \\
  \midrule
  & $G_m(a) \triangleq D_m(a) / a^m, m \in \{1, 2, \YL{3's?}\}$ \\
  \cmidrule(lr){2-2}
  & $\d\ln G_m / \d\lna$ \\
  \cmidrule(lr){2-2}
  dimensionless factors & $\Omegam(a), \YL{\OmegaK(a)}$ \\
  \cmidrule(lr){2-2}
  & $\d\ln\!H / \d\lna$ \\
  \cmidrule(lr){2-2}
  & $\Delta\lna$: time step size in $\lna$ \\
  & \YL{Now this can be an important feature with varying $a_\ic$} \\
  \bottomrule
  \end{tabular}
  \end{table}

Cosmological $N$-body simulations evolve dark matter particles under the
gravitational force.
The dynamics of the particles is governed by the Poisson equation
%
\begin{equation}
  \nabla^2 \varphi(\vx) = \frac32 \Omegam H_0^2 \delta(\vx),
\end{equation}
%
where $\Omegam$ is the matter fraction today, $H_0$ is the Hubble
constant, and the two fields $\varphi(\vx)$ and $\delta(\vx)$ are the
gravitational potential and matter density fluctuation respectively.
In Fourier space, the gravitational force field $- \vnabla \varphi$
follows
%
\begin{equation}
- i \vk \varphi(\vk) = i \frac32 \Omegam H_0^2 \,\frac{\vk}{k^2} \delta(\vk),
\end{equation}
%
we then introduce SO force sharpening through
%
\begin{equation}
\frac{i k_i}{k^2} \to \frac{i k_i}{k^2}
  f(k_i; \vvartheta) g(k_1, k_2, k_3; \vvartheta),
\end{equation}
%
where $f$ and $g$ are NNs, each approximating a nonlinear function of
the wavevector components and all possible relevant features
$\vvartheta$.
The subscript $i\in\{1,2,3\}$ denotes the 3 principal axes of the
simulation box.

Gravity obeys the $\mathrm{O}(3)$ rotational symmetry, which is however
broken by the PM solver to the O\textsubscript{h} discrete symmetry with
48 elements involving permutations and reversals of the 3 principal axes
of the simulation box.
The PM evaluation of each force component can further break the symmetry
along that axis, leaving the remaining 2 permutationally symmetric.
Therefore, we implement $f$ and $g$ to be even functions of the
wavevector components, and $g$ to be permutation equivariant in all 3
components, so that $f g$ is the most general modification to the force
kernel.
An even and real kernel in configuration space is also even in Fourier
space.
We take the absolute value of the wavevectors for the evenness, and sort
the 3 wavevector compontents to achieve the permutation equivariance.

Our target data samples are generated with the \GADGET\ simulation code
\citep{GADGET-4}.
We run 512 \GADGET\ simulations that cover a wide range of
configurations and cosmological parameters, as listed in
\autoref{tab:param}.
\YL{
Each simulation evolves $128^3$ particles from the initial time $a_\ic$
to the final time $a_\f=1$, saving 64 snapshots linearly spaced in scale
factor from $a_\ic + \gamma \Delta a$ to $1 - (1 - \gamma) \Delta a$.
We determine $a_\ic$ for each configuration depending on its cosmology
and mean particle spacing \citep{MichauxEtAl2021}, \YL{(YL elaborate why
and how here.)} which then sets the snapshot spacing $\Delta a = (1 -
a_\ic) / 64$.
To sample different alignment between the snapshots and the start/stop
times and avoid possible related numerical errors in \pmwd, we introduce
a fractional random offset $\gamma$ that varies between 0 and 1 for
different simulations.
}
The curvature $\OmegaK$ is related to the separate universe simulation
\citep{LiEtAl2014, WagnerEtAl2015}.
\YZ{We refer to this training dataset of 512 \GADGET\ simulations
sampled with Sobol sequence as \GSDATA.}
On average each \GADGET\ simulation takes roughly 2K CPU hours to run,
and the generation of the whole \GSDATA\ dataset consumes 1M CPU hours.

\YL{
In this work, we train \pmwd\ using $XXX$ time steps from $a_\ic$ to
$a_\f$, generate snapshots at the same times as in \GADGET, and compare
them in the loss function.}
In our \pmwd\ model, the snapshot at an arbitrary given time is
generated using the cubic Hermite interpolation, with more details in
\autoref{sec:snapobs}.
\YL{Although the number of time steps are fixed here, we emphasize that
our spatial optimization of course also applies for different number of
time steps, which we will explore, together with the optimal choice of
time variable, in a follow-up work on the temporal optimization.
}

\subsection{Spatial Optimization Neural Networks}

\begin{figure*}
  \centering
  \includegraphics[width=.48\columnwidth]{net_f_a1_j3091776_e3000.pdf}
  \includegraphics[width=.48\columnwidth]{net_g_a1_j3091776_e3000.pdf}
  \caption{$f(k_i; \vvartheta)$ and $g(k_1, k_2, k_3; \vvartheta)$
  functions. The solid line and shaded region indicate the median and
  the [16, 84] percentiles of the 64 Sobol simulation setups. $a=1$,
  mesh shape=1}
  \label{fig:func}
\end{figure*}

For both $f$ and $g$ NNs, we adopt multilayer perceptron (MLP, i.e.
fully connected) structures to learn the nonlinear functions.
Both $f$ and $g$ networks have $XXX$ hidden layers with $XXX\times n_{\rm
input}$ nodes for each layer, where $n_{\rm input}$ is the size of the
input layer, which is [...] and [...] for $f$ and $g$ respectively.
The rectified linear unit (ReLU) activation function is used for hidden
layers.
We further apply the exponential function to the output as a regulator
for positive output.
Being embedded in the differentiable simulation library \pmwd, the NNs
can be trained using gradient-based optimization methods as in usual
deep learning tasks.
The parameters of the NNs, which enter the force evaluation in every
time integration step, can be treated just the same as other
cosmological parameters.
The objective gradients are accumulated, based on which the parameters
are then optimized.

With discrete Fourier transform (DFT) on a regular grid of length $N_c$
and cell size $l_c$, the range of sampled 1D discrete wavenumbers is
limited by the fundamental and Nyquist values, i.e. $k_{\rm fund} =
2\pi/(l_c N_c)$ and $k_{\rm Nyquist} = \pi/l_c$, where $l_c N_c$ is
simply the box size.
Then the normzalized wavenumber $k/k_{\rm Nyquist}$ gives the same range
$[2/N_c, 1]$ for different box sizes.
For the 3D wavenumber $k=\sqrt{\sum_{i=1}^{3}k_i^2}$, the maximum
normalized value is then $\sqrt{3}$.

For $g(k_1,k_2,k_3;\vvartheta)$, which is permutationally symmetric
about $(k_1, k_2, k_3)$, we show it as a function of the wavevector
along axis, face, and body diagonal directions.

\subsection{Symbolic Expressions}

We fit the trained NNs with symbolic regression for analytic
expressions of the $f(k_i; \vvartheta)$ and $g(k_1, k_2, k_3;
\vvartheta)$ functions.

\subsection{Dark Matter Density Fields}

\begin{figure*}
  \centering
  \includegraphics[width=.48\columnwidth]{tf_snap120_3177874_e3000_mid-per.pdf}
  \includegraphics[width=.48\columnwidth]{cc_snap120_3177874_e3000_mid-per.pdf} \\
  \includegraphics[width=.48\columnwidth]{tf_test_snap120_3177874_e3000_mid-per.pdf}
  \includegraphics[width=.48\columnwidth]{cc_test_snap120_3177874_e3000_mid-per.pdf}
  \caption{The TF and CC of original PM and optimized PM-SONN simulated
  density fields with respect to that of the \GADGET.
  The solid line and shaded region indicate the median and the [16, 84]
  percentiles of different simulation configurations.
  The results are shown for the last snapshot sampled in each
  configuration, which are all around time $a\simeq 1$.
  The x-axis is the wavenumber $k$ normalized by the Nyquist wavenumber
  $k_{\rm Nyquist}$ of the particle grid, which is therefore the same
  for different box sizes.
  \textit{Upper}: results for the training dataset.
  \textit{Lower}: results for the test dataset.
  }
  \label{fig:tfcc}
\end{figure*}

Improvements of SO on the dark matter density fields are the first
things to check.
We construct these fields by scattering the particles onto a mesh.
To compare two fields $\hat f$ and $f$, we introduce for each mode the
square root ratio of the auto power spectrum
%
\begin{equation}
T(\vk) \triangleq
\sqrt{\frac{|\hat f(\vk)|^2}{|f(\vk)|^2}},
\label{T}
\end{equation}
%
which is referred to as the transfer function (TF), and the correlation
coefficient (CC) defined with the cross power spectrum
%
\begin{equation}
r(\vk) \triangleq
\frac{\Re \bigl[ \hat f(\vk) f^*(\vk) \bigr]}
     {\sqrt{|\hat f(\vk)|^2 |f(\vk)|^2}}.
\label{r}
\end{equation}
%
The two fields are identical when both $T$ and $r$ are unity.
The PM method is accurate on scales much larger than the mesh cell size,
therefore we expect $T$ and $r$ to be very close to being unity both
before and after the SO with force sharpening.

In \autoref{fig:tfcc}, we show the results of $T$ and $r$ for both
training and test data samples.
To show different box sizes, we normalize the wavenumber by the Nyquist
wavenumber $k_{\rm Nyquist} = \pi / l_p$ of the particle grid, where
$l_p$ is the particle spacing.

comparison of PM, PM with SO NN (PM-SONN), PM with SO SR (PM-SOSR)

\begin{figure*}
  \centering
  \includegraphics[width=.8\columnwidth]{denslab_0_120_e300.pdf}
  \caption{Projected dark matter density fields of \GADGET, \pmwd\ without and
  with force sharpening.
  Sobol 0 snap 120}
  \label{fig:denslab}
\end{figure*}

% density field with zoom in (with halos painted as circles?)
The effect of force sharpening can also be visually shown by looking
directly at the density field.

\subsection{Dark Matter Halos}

\subsection{Benchmarks}
We compare the computational cost of \GADGET, \pmwd\ with NNs, and
\pmwd\ with SR.

We time the computation of force sharpening for one step using NNs and
SR respectively.

\vspace{1em}
\section{Discussion}
% Discussion does not contain subheadings

\vspace{1em}
\section{Methods}

\vspace{1em}
\subsection{Differentiable Simulation}
\label{sec:pmwd}

The SO using gradient descent training discussed in this work is built
in the framework of the PM $N$-body library \pmwd\ \citep{Li2022b},
which is an implementation of the differentiable cosmological simulation
with the adjoint method \citep{Li2022a}.

Here we briefly recap the equations that are closely related to our
extensions of the original framework, and introduce the notations of the
variables.
Following \cite{Li2022a}, we denote the particles' state with $\vz =
(\vx, \vp)^\intercal$, where $\vx$ and $\vp$ are the positions and
momenta of the particles, and their adjoint variables with $\vlambda =
(\vxi, \vpi)$.
We refer to the quantities that are directly related to the final
objective function $\cJ$ as the observable $\vo$.
In this work, our observables include multiple snapshots at different
times.
For the observation operator $O_{i}^{i-1}$ in the adjoint equations
(check Eq.(49) in \cite{Li2022a}) from time step $a_i$ back to
$a_{i-1}$, the adjoint variables are updated through
\begin{align}
  O_i^{i-1}: \qquad \vxi_{i-1} &:= \vxi_{i-1} +
              \frac{\p\cJ}{\p\vo}\frac{\p\vo}{\p \vx_{i-1}}, \nonumber\\
  \vpi_{i-1} &:= \vpi_{i-1} +
                          \frac{\p\cJ}{\p\vo}\frac{\p\vo}{\p \vp_{i-1}},
\end{align}
where the derivatives of $\vo$ with respect to $\vx_{i-1}$ or
$\vp_{i-1}$ would be non-zero only if $\vx_{i-1}$ or $\vp_{i-1}$ was
used in the generation of any snapshots in $\vo$.
The initial conditions of these adjoint variables are given by
\begin{align}
  \vxi_n &= \frac{\p\cJ}{\p\vo}\frac{\p\vo}{\p\vx_n}, \nonumber\\
  \vpi_n &= \frac{\p\cJ}{\p\vo}\frac{\p\vo}{\p\vp_n},
\end{align}
which are zero if no snapshot in $\vo$ depends on $(\vx_n, \vp_n)$, the
state of particles at the final time step.
The derivative of the objective function with respect to the observable,
i.e. $\p\cJ/\p\vo$, depends on our construction of the loss function on
the snapshots (\autoref{sec:loss}).
The VJPs of the observation operation, i.e. $\p\vo/\p\vx_n$ and
$\p\vo/\p\vp_n$, mainly involves the interpolation function
(\autoref{sec:snapobs}).
Both can be conveniently computed by AD.

\vspace{1em}
\subsection{Snapshot Observable}
\label{sec:snapobs}

The $N$-body time integration is carried out along discrete time steps
in terms of the scale factor $a$, and the state of particles at any
given time can be achieved using interpolation.
Given the particle displacements $\vs$ and velocities\footnote{This
velocity is related to the canonical momentum $\vp$ through $\vv =
\vp/(a^3H)$.} $\vv := \d\vs/\d a$ at two consecutive time steps $a_i$ and
$a_{i+1}$, we get the snapshot of particles in between at $a$ using the
cubic Hermite interpolation.
The interpolated particle displacements are given by
\begin{align}
  \vs(a) =\ &h_{00}(\alpha)\vs_i + h_{10}(\alpha)(a_{i+1} - a_i)\vv_i + \nonumber\\
           &h_{01}(\alpha)\vs_{i+1} + h_{11}(\alpha)(a_{i+1} - a_i)\vv_{i+1},
\end{align}
where $\alpha := (a - a_i)/(a_{i+1} - a_i)$ and the Hermite basis
functions read
\begin{align}
  h_{00}(\alpha) &= 2\alpha^3 - 3\alpha^2 + 1, \nonumber\\
  h_{10}(\alpha) &= \alpha^3 - 2\alpha^2 + \alpha, \nonumber\\
  h_{01}(\alpha) &= -2\alpha^3 + 3\alpha^2, \nonumber\\
  h_{11}(\alpha) &= \alpha^3 - \alpha^2.
\end{align}
The particle velocities are then given by the derivative $\vv(a) =
\d\vs(a) / \d a$.
The interpolation is a linear combination of the particle snapshots at
two consecutive time steps, which therefore can be done in either two
steps or just one step by carrying the particles from the previous step.
We find the two-steps scheme more convenient for the backward evolution
of the adjoint variables.
The interpolated snapshots at different times make the observable in
this work, and the VJPs of the interpolation functions are computed by
AD and put into the adjoint equations as mentioned in
\autoref{sec:pmwd}.

\vspace{1em}
\subsection{Configurations and Data Samples}

\begin{figure*}
  \centering
  \includegraphics[width=0.9\linewidth]{sobol.pdf}
  \caption{Randomized Quasi-Monte Carlo (RQMC) configuration with
    scrambled Sobol sequence of 512 points in 9D.
    Lower triangular panels show the 2D projections and the diagonal
    panels are the 1D cumulative histograms.
    From left to right (top to bottom), we use each dimension of the
    sample to scale the parameters as ordered in \autoref{tab:param}.
    We use the \texttt{scipy.stats.qmc} package \citep{SciPy} to
    generate the Sobol sequence \citep{Sobol1967}, which uses the
    direction number from \citet{JoeKuo2008} and the Owen scrambling
    \citep{Owen1998}.
    We search among 65536 scrambling seeds to minimize the mixture
    discrepancy (a uniformity measure) proposed in \citet{Zhou2013MD}.
  }
  \label{fig:sobol}
\end{figure*}

We configure \GADGET\ with high force and time integration accuracy
settings.
For gravitational force, we choose pure FMM with $p=5$ and opening angle
0.4.
%FIXME list other from Config.sh and param.txt

We sample 512 data points in the 9D space of configurations and
cosmological parameters using randomized quasi-Monte Carlo (RQMC) method
with scrambled Sobol sequence (\autoref{fig:sobol}).
We use the \texttt{scipy.stats.qmc} package \citep{SciPy} to generate
the Sobol sequence \citep{Sobol1967}, which uses the direction number
from \citet{JoeKuo2008} and the Owen scrambling \citep{Owen1998}.
We search among 65536 scrambling seeds to minimize the mixture
discrepancy (a uniformity measure) proposed in \citet{Zhou2013MD}.

In addition, we sample 9 extra data points for the interpolation test of
the trained model to different configurations and cosmological
parameters.

Small numerical differences in initial conditions (ICs) could become
more significant with the $N$-body evolution.
We verified that the IC generation on GPUs is deterministic though
different from those on CPUs, therefore we generate ICs for both the
training data and the \pmwd\ model on GPUs.

\vspace{1em}
\subsection{Input Features}
\label{sec:features}

We design the input features of the NNs in a way such that they could
properly fit and interpolate given our wide coverage of the
configurations and cosmological parameters.

Multiple scales and dimensionless factors could possibly affect force
sharpening, which are denoted as $\vvartheta$ and enter the NNs of $f(k_i;
\vvartheta)$ and $g(k_1, k_2, k_3; \vvartheta)$ as input features.
A list of these features is summarized in \autoref{tab:feat}, and here we
present relevant definations and more details.

Below are some definations related to the typical scales.
\begin{itemize}
\item The root-mean-square (RMS) linear displacement is given by
%
\begin{equation}
R_\mathrm{d} = \int \frac{k P_\linear(k)}{6\pi^2} \d\lnk,
\end{equation}
%
where $P_\linear(k)$ is the linear power spectrum.
\item The dimensionless linear power spectrum is defined as
%
\begin{equation}
\Delta^2_\linear(k) \triangleq \frac{k^3 P_\linear(k)}{2 \pi^2},
\end{equation}
%
based on which we define a nonlinear scale $R_P$ through
$\Delta_\linear^2(R_P^{-1}) \triangleq 1$.
\item The variance of linear matter overdensity in a top-hat window is given by
%
\begin{equation}
\sigma_\tophat^2(R) = \int_0^\infty \frac{\d k}k
  \Delta_\linear^2(k) W_\tophat^2(kR),
\end{equation}
%
where $W_\tophat(kR) = 3[\sin(kR) - kR\cos(kR)] / (kR)^3$ is the
Fourier transform of the top-hat window.
Then we have $R_\tophat$: $\sigma_\tophat^2(R_\tophat) \triangleq 1$.
\item Likewise $\sigma_\gauss^2(R)$ is defined with a Gaussian window
$W_\gauss(kR) = e^{-(kR)^2/2}$, and $R_\gauss$: $\sigma_\gauss^2(R_\gauss)
\triangleq 1$.
\end{itemize}
These nonlinear scales are functions of time since the linear power
spectrum evolves with time through the growth factor $D(a)$, we also
take the time derivatives of these scales as input features.
Other relevant scales include the particle spacing in Lagrangian space,
the cell size of the chosen mesh grid for force evaluation, and the
softening length in the target \GADGET\ simulations.

Relevant scale-independent dimensionless factors are
\begin{itemize}
\item $G_m(a) \triangleq D_m(a) / a^m$: linear growth given in
  suppression factors for $m \in \{1, 2\}$;
\item $\d\ln G_m / \d\lna$: with $m=1$ case related to the growth rate
  $f \triangleq \d\ln D_1 / \d\lna$;
\item $\Omegam(a)$: matter density parameter at $a$;
\item $\d\ln\!H / \d\lna$: time derivative of Hubble expansion;
\item $\Delta\lna$: time step size in $\lna$.
\end{itemize}

Dimensionless combinations are formed by multiplying the wavevector
components with the scales, i.e.,
%
\begin{equation}
f(k_i; \vvartheta) = f(k_iR_d, k_iR_P, \cdots; G_m, \Omegam, \cdots),
\end{equation}
%
\begin{equation}
g(k_1, k_2, k_3; \vvartheta) = g(\{k_jR_d, k_jR_P, \cdots\}_{j=1,2,3};
                                 G_m, \Omegam, \cdots).
\end{equation}
%

\vspace{1em}
\subsection{Loss Function}
\label{sec:loss}

The loss function is constructed by comparing the simulated field $\hat
f$ with the target field $f$.
In this work, we consider two fields in our loss function, the
Lagrangian particle displacement field $\vs(\vq)$ and the Eulerian
density field $\delta(\vx)$.

For the displacement field, the mean squared error (MSE) loss has been
proved effective in many previous works
\citep[e.g.,][]{HeEtAl2019, LiEtAl2021}:
%
\begin{equation}
\cL_\vs = \ln \frac{\sum_\vq \bigl[ \hat\vs(\vq) - \vs(\vq) \bigr]^2}
                   {\sum_{\vq'} s(\vq')^2},
\end{equation}
%
in which a particle originating from its Lagrangian position $\vq$ is
displaced by $\vs$ in the \GADGET\ snapshot and by $\hat\vs$ in the
prediction of \pmwd\ model in training.
Note that we have further taken the logarithm to combine it with the
other loss(es).
Even though the logarithm should be enough for adaptive optimizer and
fixed simulation configuration, this may create problem in our training
on all configurations at the same time.
So normalizations inside the logarithm is added above and below.
% Say that only s loss is not enough even though in principle it is,
% and cite DrewEtAl x2 too

However, in the Eulerian space, an MSE loss is dominated by the small
scale modes.
In this work we choose uniform weights in the scale space on the
relative MSE.
We bin the modes logarithmically by \nicefrac13 octave in scale, with
the $i$-th $k$-bin $K_i \triangleq [k_i, k_{i+1})$, where $k_i =
2^{1+i/3} \pi / L$.
We introduce a loss function that sums up its MSE relative to the target
field power in logarithmic bins:
%
\begin{equation}
  \cL_f = \frac1{N_K} \sum_{i} \ln \biggl[
  \frac{\sum_{\vk \in K_i} \bigl| \hat f(\vk) - f(\vk) \bigr|^2}
       {\sum_{\vk' \in K_i} | f(\vk')|^2} + \epsilon \biggr],
\label{loss_power_ln}
\end{equation}
%
where $N_K$ is the number of $k$-bins, and $\epsilon$ sets the relative
threshold tolerance, below which we consider $\hat f$ in each $K$
accurate enough and the gradients from that bin less important.
% Under the mild assumptions that $T$ \eqref{T} and $r$ \eqref{r} are smooth
% and isotropic, we can show
% %
% \begin{equation}
% \cL_\delta \propto \int \!\d\ln\!k\, \ln
% \Bigl\{ \bigl[ 1 - T(\vk) \bigr]^2
%   + 2 T(\vk) \bigl[ 1 - r(\vk) \bigr] \Bigr\} \geq -\infty,
% \end{equation}
% %
% replacing
% %
% \begin{equation}
% \cL_\delta \propto \int \!\d\ln\!k\, w(\vk)
% \Bigl\{ \bigl[ 1 - T(\vk) \bigr]^2
%   + 2 T(\vk) \bigl[ 1 - r(\vk) \bigr] \Bigr\} \geq 0,
% \end{equation}
% %
% that only vanishes when both $T$ and $r$ are unity for all modes, i.e.,
% when $\hat\delta$ is identical to $\delta$.
% Above we choose the binned version instead for the robustness against
% the noisy modewise $|\delta(\vk)|^2$ in the denominator.
% The derivation here is similar to that in \citet{HeEtAl2019} on the MSE
% loss, while our modification is important to balance among all scales so
% that the neural network does not focus too much on either high or low
% $k$'s.

\YZ{In this work, we use the loss function in \eqref{loss_power_ln} for
both $\vs(\vq)$ and $\delta(\vx)$.}
We combine the two fields to form the total loss for a single snapshot,
and further sum over all snapshots in a single forward simulation
%
\begin{equation}
\cJ = \sum_{a_i} \left[\cL_{\vs(a_i)} + \cL_{\delta(a_i)}\right].
\end{equation}
%

\vspace{1em}
\subsection{Training of Neural Networks}

We perform distributed data-parallel training using GPU devices hosted
on multiple computing nodes.
Each process is binded to a single GPU device, and the training data
samples are uniformly distributed to the processes.
In total, we use XXX \GPU\ GPUs, each accompanied with XXX CPU cores
and XXX GB CPU memory mainly for preloading the training data.
The order of data samples within each process is shuffled across
training epochs.
In each training step, a data sample is fed into the GPU from the CPU
memory where the preloading took place to save the repeated reading time
from the disk.

For each training step, we configure and run \pmwd\ with the same
initial coniditions, cosmological parameters and other common
configurations as the training data sample.
As discussed in \autoref{sec:loss}, the loss in a single simulation sums
over all snapshots.
Our batch size is XXX, i.e. the number of GPU devices, since one
simulation is run on a single GPU and compared to the corresponding
training data sample for loss.
Then the loss and gradients are collected and averaged across all GPU
devices, with which the optimizer is updated.
We use the Adam \citep{Adam} optimizer.

We initialize the parameters of NNs in a way such that the output is
very close to one, i.e. no SO.
This is achieved by setting the kernel of the output layer to be very
small random values and the bias to be zero, whose return is then around
one after the exponential function.
The parameters are also initialized to be the same for all devices, and
hence remain the same during training given the same averaged gradients.

% We use Optuna \citep{optuna_2019} to search for the optimal
% hyperparameters.

% We empoly a learning rate scheduler that reduces the learning rate on
% plateau where the loss function stops decreasing.
% The initial learning rate is set at [...] with a reduction factor of
% [...] after [...] epochs of no decreasing on loss.

Same as the \pmwd\ library, this work is also implemented using JAX
\citep{JAX}.
We implement the NNs with \texttt{Flax} \citep{flax2020github}, and use
the optimizers provided in \texttt{Optax}, part of the DeepMind JAX
Ecosystem \citep{deepmind2020jax}.

\vspace{1em}
\subsection{Symbolic Regression Fitting}

To perform further analyses of the trained NNs, including feature
importance analysis and symbolic regression fitting, we need to generate
samples of (input, output) data.
The input features of the NNs are constructed based on the RQMC
parameters in \autoref{tab:param}, the wavenumber $k$, and the scale
factor $a$.
Again, we use Sobol sequence to sample data points in this high
dimensional space.

We use the random forest regressor for feature selection using
impurity-based importance.
We use the \texttt{scikit-learn} \citep{scikit-learn} code for this.
% more details on the hyperparams in random forest
\YZ{Maybe we can add another column in \autoref{tab:feat} for feature
importance?}

The selected important features are then used for SR.
We use the \texttt{PySR} \citep{PySR} SR code to fit for analytic
function expressions of the trained SO neural networks.

\YL{It is interesting to symbolically regress $f$, $k_i f$, and $f g$
etc to see which one gives simpler expression, given the possible
different rankings by complexity.
For example, $k_i f(k_i)$ may have a simpler form as a nonlinear odd
function.}

\vspace{1em}
\textit{\large Acknowledgements:}
We thank Qiufan Lin, Paulo Montero-Camacho, and Yang Wang for helpful
discussions.
YZ and YL were supported by The Major Key Project of PCL.
YZ was also supported by the China Postdoctoral Science Foundation under
award number 2023M731831.
The Flatiron Institute is supported by the Simons Foundation.

\vspace{1em}
\textit{\large Data Availability:}
We make the \GSDATA\ simulation data set public at ...

\vspace{1em}
\textit{\large Code Availability:}
\pmwd\ \citep{Li2022b} is open-source on GitHub
\href{https://github.com/eelregit/pmwd}{\faGithub}, including the source
files and scripts of this paper
\href{https://github.com/eelregit/pmwd/tree/master/docs/papers/sto}{\faFile}.
\GADGET\ \citep{GADGET-4} can be obtained through a public
git-repository \href{http://gitlab.mpcdf.mpg.de/vrs/gadget4}{\faGitlab},
hosted by a gitlab server at the Max-Planck Computing and Data Facility
(MPCDF).

\bibliographystyle{aasjournal}
\bibliography{sto}

\vspace{1em}
\appendix
% supplementary information of Nat. Mach. Intell.

\vspace{1em}
\section{Tests of Snapshots Observation in Adjoint Method}
As discussed in the main text, our objective loss function is defined on
multiple snapshots at given times, which are constructed through
interpolation during the observation operation.
Here we perform tests on the gradients of NN parameters to check our
implementation of the snapshots observation inside the adjoint method of
the $N$-body time integration.
The gradients given by the adjoint method are compared against those
given by the AD where we disable the custom VJP of the $N$-body
function, see
\autoref{fig:nn_grads_cmp}.

\begin{figure*}
  \centering
  \includegraphics[width=0.5\linewidth]{nn_grads_cmp.pdf}
  \caption{A comparison of adjoint gradients of all parameters in the
  $f$ and $g$ NNs to those by AD.
  This test runs a \pmwd\ simulation of $128^3$ particles, with a
  $128^3$ mesh and 15 time steps.
  The size and time steps of this test case is mainly limited by the
  memory cost of the AD method.
  The two NNs $f$ and $g$ both have two hidden layers of size $[n, n]$,
  where $n$ is the size of the input layer.
  The loss function involves 3 snapshots at $a=(0.6, 0.8, 1)$.
  }
  \label{fig:nn_grads_cmp}
\end{figure*}

%\listofchanges

\section*{TODO}
\begin{itemize}
\item (data; YL) Incorporate 3lpt branch; testing; find a\_start for
  different mass resolutions; probably replace linear snapshot spacing
  according to the nonlinearity (estimate (with PM till convergence) the
  std acc over the std vel) of each realization.
\item (train, test) 3 sets of grid for ptcl, force mesh, and loss. Add
  random offset in loss grid (see \autoref{tab:param}) to remove the
  position dependence in loss; likewise for force mesh? how about only
  training (for lower loss), only ``inference'' (for higher accuracy),
  or both? the random force mesh offset can also be replaced by a
  structured scheme, e.g., alternating between 2 interlocking grids
\item (train, test) Non-integer multiple mesh shapes: for uniform
  particle grid, test (with double precision on CPUs) whether they start
  moving by artifact forces on cell scales, in 3 cases: no
  deconvolution, with CIC deconvolution, with SO.
\item (test) Interpolation test: compare SO DM and halo fields and
  summary stats with those of raw pmwd, deconvolved pmwd, and \GADGET,
  on the last 9 test cases and noncubic box.
\item (test) Extrapolation test: demonstrate generalization capability,
  compare extrap PySR, extrap SO, raw pmwd, deconv pmwd, and \GADGET.
\item Fix cross references and other journal quirks as a patch
\end{itemize}

\end{document}